### 2020-05-28 serial.tex
almost done

Generation intervals and serial intervals are critical quantities for characterizing outbreak dynamics.
Generation intervals characterize the time between infection and transmission; whereas serial intervals characterize the time between the onset of symptoms in a chain of transmission.
They are often used interchangeably, leading to misunderstanding of how to use either generation or serial intervals to establish a link between the epidemic growth rate $r$ and the reproduction number \RR.
On one hand, generation intervals provide a mechanistic link between $r$ and \RR but are harder to measure via contact tracing.
On the other hand, serial intervals are easier to measure from contact tracing but recent studies suggest that the two intervals give different estimates of \RR from $r$.
We present a general framework for characterizing \emph{any} epidemiological delays based on cohorts (i.e., a group of individuals that share the same event time, such as symptom onset) and show that \emph{forward-looking} serial intervals, which correctly estimates \RR from $r$, are not the same as ``intrinsic'' serial intervals, but instead change with $r$.
We provide a heuristic method for addressing potential biases that can arise from not accounting for changes in serial intervals across cohorts and apply the method to estimating \RR for the COVID-19 outbreak in China using serial interval data --- our analysis shows that using incorrectly defined serial intervals can severely bias the estimate.
This study demonstrates the importance of early epidemiological investigation through contact tracing and provides a rationale for reassessing generation intervals, serial intervals, and \RR\ estimates, for COVID-19.

@@

The reproduction number \RR is one of the most important characteristics of an emerging epidemic, including the current pandemic of coronavirus disease 2019 (COVID-19) \citep{majumder2020early}.
The reproduction number is defined as the average number of secondary cases caused by a primary case;
the value in a fully susceptible population --- referred to as the basic reproduction number \Ro --- allows us to predict the extent to which an infection will spread in the population and the amount of intervention necessary to eliminate it \citep{anderson1991infectious}.
Near the beginning of an outbreak, \RR is often estimated from the observed exponential growth rate using generation- and serial-interval distributions (e.g., \cite{wallinga2007generation, fraser2009pandemic, hampson2009transmission, chunara2012social, chowell2014west, du2020serial, jung2020real}).

@@

The time delays between two epidemiological events can be defined either within an infected individual (e.g., incubation period: infection and symptom onset of an individual) or between infected individuals (e.g., serial interval: symptom onsets of an infector and an infectee).
We can further divide these events into \emph{primary} and \emph{secondary} events.
When we measure an epidemiological time delay within an infected individual (e.g., the incubation period), the primary event is the event that usually occurs before the secondary event ---
most epidemiological events that can be observed within an individual have clear direction (e.g., infection and onset of symptoms) but some may not (e.g., onset of infectiousness and onset of symptoms).
When we measure an epidemiological time delay between infected individuals (e.g., the serial interval), 
the primary and secondary events are defined in terms of the direction of transmission:
the primary event refers to the event that occurs in the infector. 
Again, some of these delays are always positive (the infector is always infected before the infectee) and some are not (it is possible for the infectee to develop symptoms before the infector \citep{he2020temporal}).

@@

Both forward and backward perspectives
must yield identical \emph{measurement} (e.g., the length of the incubation
period of a given individual is the same whether measured forward from
the time of infection or backward from the time of symptom onset).
Consequently, if we restrict attention to individuals with a given
incubation period, say $\tau$, and we count how many became infected at time $\psymp$
or how many first showed symptoms at time $\ssymp=\psymp+\tau$ then we must obtain
the same number. In general, no matter how delays are distributed, if
$\mathcal P$ and $\mathcal S$ represent the sizes of primary and
secondary cohorts then we must have
\begin{equation}
\psize(\psymp) f_\psymp(\tau) = \ssize(\ssymp) b_\ssymp(\tau) \,.
\label{eq:match}
\end{equation}
Substituting $\psymp=\ssymp-\tau$, it follows that
\begin{equation}
b_\ssymp(\tau)= \frac{\psize(\ssymp-\tau) f_{\ssymp-\tau} (\tau)}{\ssize(\ssymp)}\,.
\label{eq:backward}
\end{equation}
If we are thinking of realized incubation periods, then the left hand side of
this equation is the probability density that an individual
who became symptomatic at time $\ssymp$ had an incubation period of
length $\tau$. From the right hand side, we see that this probability
density depends on how many individuals became infected at time
$\ssymp-\tau$ [$\psize(\ssymp-\tau)$], how many individuals became
symptomatic at time $\ssymp$ [$\ssize(\ssymp)$], and the probability density
that someone who was infected at time $\ssymp-\tau$ became symptomatic at
time $\ssymp$ [$f_{\ssymp-\tau}(\tau)$]. In general, Eq.~\eqref{eq:backward}
shows that the backward delay distribution at reference time $\ssymp$
[$b_\ssymp(\tau)$] depends not only on the secondary cohort size at time
$\ssymp$ [$\ssize(\ssymp)$], but also on the primary cohort size at the
earlier time $\ssymp-\tau$ [$\psize(\ssymp-\tau)$] and the forward delay
distribution at that earlier time [$f_{\ssymp-\tau}(\tau)$].

Several different mechanisms drive the changes in forward and backward delay distributions over time.
Typically, within-individual forward delay distributions are not affected by epidemic dynamics.
Some realized distributions, like incubation period distributions, are expected to be equivalent to their intrinsic distributions and remain invariant over the course of an epidemic.
Other realized distributions, like the distribution of time from symptom onset to testing, may change over the course of an epidemic due to changes in public health policies.
Between-individual forward delay distributions, such as generation- or serial-interval distributions, depend on epidemic dynamics;
for example, forward generation intervals contract during the outbreak because infected individuals are less likely to infect others later in the epidemic due to factors that drive time-dependent decreases in transmission such as susceptible depletion, behavioral change, and interventions \citep{champredon2015intrinsic}.

\eref{backward} suggests that backward delay distributions change over time even if their corresponding forward delay distribution does not change.
Backward delay distributions depend on changes in the primary cohort size over time due to conditionality of observations:
Conditioning on individuals whose secondary events have occurred at the same time means that the distribution is truncated so that we tend to observe shorter (or longer) inter-event delays when incidence is  the cohort size has been increasing (or decreasing) over time.
In particular, when the incidence is growing exponentially, we can calculate the amount of bias exactly.
Assuming that the forward delay distribution remains constant during the exponential growth phase ($f_p(\tau) \approx f_0(\tau)$), we can substitute $\psize(t) = \psize(0) \exp(rt)$ in \eref{backward} to obtain:
\begin{equation}
b_0(\tau) = \frac{\exp(-r\tau) f_0(\tau)}{\left[\int_{-\infty}^{\infty} \exp(-r\tau') f_0(\tau') \dtau'\right]},
\label{eq:backexp}
\end{equation}
where $r$ is the exponential growth rate and the expression in square brackets is a constant since it does not depend on $\tau$.
For faster epidemics (higher $r$), the backward delay distribution will have a shorter mean.
Changes in the backward delay distribution are also by definition affected by the changes in the forward delay distribution (\eref{backward}).
Therefore, the backward delay distribution will generally differ from the forward delay distribution even if we are measuring time delays that are intrinsic to the life history of a disease (e.g., the incubation period).
These ideas apply to all epidemiological delay distributions and generalize the work by \cite{champredon2015intrinsic} who compared forward and backward generation-interval distributions to describe the realized generation intervals from the perspective of an infector and an infectee, respectively, as well as the work by \cite{britton2019estimation} who showed that \eref{backexp} holds for the backward generation-interval distributions.

@@

We illustrate changes in the forward and backward serial intervals over the course of an epidemic by applying it to a specific example of an epidemic model. 
We model disease spread with a renewal-equation model \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral, roberts2007model, champredon2018equivalence}.
Ignoring births and deaths, changes in the proportion of susceptible individuals $S(t)$ and incidence of infection $i(t)$ can be written as:
\begin{align}
\frac{\mathrm{d}S}{\mathrm{d}t} &= - i(t)\nonumber\\
i(t) &= \Ro S(t) \int_0^\infty i(t-\tau) \gdist(\tau) \dtau,
\label{eq:renewal}
\end{align}
where \Ro is the basic reproduction number, and $\gdist(\tau)$ is the intrinsic generation-interval distribution (i.e., the forward generation-interval distribution of a primary case in a population where the number of susceptibles remains constant).
Then, the forward generation-interval for a cohort of individuals that were infected at time $\psymp$ follows \citep{champredon2015intrinsic}:
\begin{equation}
\gdist_\psymp (\tau) = \frac{ \gdist(\tau) S(\psymp + \tau)}{\int_0^\infty \gdist(\tau') S(\psymp + \tau') \dtau'},
\end{equation}
\djde{I don't like $\propto$.  It is always clearer to state the
  constant of proportionality.}
which allows us to separate the joint probability distribution $h_\psymp$ of the forward incubation period and the forward generation-interval distribution as a product of the proportion of susceptible individuals $S$ and the joint probability distribution $h$ of the forward incubation period and the intrinsic generation intervals:
\begin{equation}
h_\psymp (x, \tau) = \frac{h(x, \tau) S(\psymp + \tau)}{\int_0^\infty \int_0^\infty h(x', \tau') S(\psymp + \tau') \dtau' \dx'}.
\end{equation}
We further assume that the forward incubation period distribution does
not vary across cohorts over the course of an epidemic, as it
represents the life history of a disease; we denote it as $\idist$.
Then, we have:
\begin{align}
\idist(x) &= \int_0^\infty h(x, \tau) \dtau\nonumber\\
\gdist(\tau) &= \int_0^\infty h(x, \tau) \dx
\label{eq:marginal}
\end{align}
Finally, the case reproduction for this model is defined as follows:
\begin{equation}
\Rc(t) = \Ro \int_0^\infty \gdist(\tau) S(t+\tau) \dtau.
\end{equation}
The forward and backward serial-interval distributions are then calculated by substituting these quantities into \eref{fserial} and \eref{bserial}.
We use this framework to illustrate how the realized epidemiological time distributions vary over the course of an epidemic and depend on the perspective (i.e., forward vs. backward).

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.95\textwidth]{forward.pdf}
\caption{
\textbf{Epidemiological dynamics and changes in mean forward and backward delay distributions.}
(A) Daily incidence over time.
(B--D) Changes in the mean forward incubation period, generation interval, and serial interval.
(E--G) Changes in the mean backward incubation period, generation interval, and serial interval.
Black lines represent the results of a deterministic simulation.
Orange lines (A) represent the results of 10 stochastic simulations.
Orange points (B-G) represent the average of 10 stochastic simulations.
Dashed lines represent the mean initial forward delay.
Intrinsic incubation periods and intrinsic generation-intervals are assumed to be independent of each other.
See Table 1 for parameter values.
\bmb{I added 50\% quantiles to one of the plots.  You can decide whether you think it's too ugly OR extend it to the other plots\ldots}
}
\label{fig:epi}
\end{center}
\end{figure}

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{serial_analysis.pdf}
\caption{
\textbf{Observed serial intervals of COVID-19 and cohort-averaged estimates of \RR.}
(A--B) forward and backward serial intervals over time.
Points represent the means. 
Vertical error bars represent the 95\% equi-tailed quantiles.
Solid lines represent the estimated locally estimated scatterplot smoothing (LOESS) fits.
The dashed line represents the maximum and minimum observable delays across the range of \bmb{???}.
(C) Cohort-averaged estimates of \Ro assuming doubling period of 6 and 8 days \citep{li2020early, wu2020nowcasting}.
Ribbons represent the associated 95\% bootstrap confidence intervals.
The data were taken from Supplementary Materials of \cite{du2020serial}.
\bmb{I think this would be slightly more readable if you nudged the min/max labels up slightly so they were clear of the dashed lines}
}
\label{fig:du}
\end{figure}

@@

Since the right hand side of \eref{ptarget} is also a
Laplace transform of $f_0=\hat{z}*\idist$, we can express it as the product
of Laplace transforms of $\hat{z}$ and $\idist$:
\begin{equation}
\int_{-\infty}^\infty \exp(-r\tau) f_{0}(\tau) \mathrm{d} \tau = \int_{-\infty}^\infty \exp(-r\tau) \hat{z}(\tau)\, \mathrm{d} \tau \int_{0}^\infty \exp(-r\tau) \idist(\tau) \dtau\,.
\label{eq:newtarget}
\end{equation}
In order to derive an expression for a Laplace transform of $\hat{z}$, we have to first derive an analytical expression for $\int_{-\infty}^\infty z(x) \dx$. By changing the order of integration, we have:
\begin{align}
\int_{-\infty}^\infty z(\sinf) \mathrm{d}\sinf &= \int_{-\infty}^\infty \int_{\max{(0,-\sinf)}}^{\infty} \exp(-r\pinf) h(\pinf, \sinf + \pinf)\, \mathrm{d}\pinf \,\mathrm{d}\sinf\nonumber\,,\\
&= \int_{0}^\infty \int_{-\pinf}^\infty \exp(- r \pinf) h(\pinf, \sinf+\pinf)\, \mathrm{d}\sinf\,\mathrm{d} \pinf\,.
\end{align}
Since $\idist$ is a marginal probability distribution of $h$, it follows that:
\begin{equation}
\int_{-\infty}^\infty z(\sinf) \mathrm{d}\sinf = \int_{0}^\infty \exp(- r \pinf) \idist(\pinf)\, \mathrm{d}\pinf\,.
\end{equation}
Then, we have:
\begin{equation}
\hat{z}(\sinf) = \frac{\int_{\max{(0,-\sinf)}}^{\infty} \exp(-r\pinf) h(\pinf, \sinf + \pinf)\, \mathrm{d}\pinf}{\int_{0}^\infty \exp(- r \pinf) \idist(\pinf)\, \mathrm{d}\pinf}\,.
\end{equation}
Substituting the expression into \eref{newtarget}, we have:
\begin{equation}
\int_{-\infty}^\infty \exp(-r\tau) f_{0}(\tau)\dtau = \int_{-\infty}^\infty \exp(-r\tau) \int_{\max{(0,-\tau)}}^{\infty} \exp(-r\pinf) h(\pinf, \tau + \pinf)\, \mathrm{d}\pinf \dtau\,.
\label{eq:newtarget2}
\end{equation}

Recall that $g$ is also a marginal probability distribution of $h$:
\begin{equation}
g(\tau) = \int_0^\infty h(\pinf, \tau) \,\mathrm{d} \pinf\,.
\end{equation}
We can then substitute $\sigma = \pinf + \tau$ into \eref{newtarget2} and apply change of variables to obtain:
\begin{align}
&\int_{-\infty}^\infty \exp(-r\tau) f_{0}(\tau)\dtau\\
&=\int_{-\infty}^{\infty} \exp(-r\tau) \int_{\max(0, -\tau)}^\infty \exp(- r \pinf) h(\pinf, \tau+\pinf) \,\mathrm{d} \pinf \dtau\\
&=\int_{0}^{\infty} \int_{0}^\infty \exp(- r \sigma) h(\pinf, \sigma)\, \mathrm{d} \pinf\, \mathrm{d}\sigma\\
&=\int_{0}^{\infty} \exp(-r\tau) g(\tau) \dtau =\frac{1}{\Ro}
\end{align}
Therefore, the initial forward serial-interval distribution and the intrinsic generation-interval distribution give the same estimates of \Ro from $r$.\qed

>>>>

Generation intervals and serial intervals are critical quantities for characterizing outbreak dynamics.
Generation intervals characterize the time between infection and transmission; whereas serial intervals characterize the time between the onset of symptoms in a chain of transmission.
They are often used interchangeably, leading to ongoing misunderstanding of how to use either generation or serial intervals to establish a link between the epidemic growth rate $r$ and the reproduction number \RR.
Generation intervals provide a mechanistic link between $r$ and \RR but are harder to measure via contact tracing.
While serial intervals are easier to measure from contact tracing, recent studies suggest that the two intervals give different estimates of \RR from $r$.
We present a general framework for characterizing \emph{any} epidemiological delays based on cohorts (i.e., a group of individuals that share the same event time, such as symptom onset) and show that \emph{forward-looking} serial intervals, which correctly estimates \RR from $r$, are not the same as ``intrinsic'' serial intervals, but instead change with $r$.
We provide a heuristic method for addressing potential biases that can arise from not accounting for changes in serial intervals across cohorts and apply the method to estimating \RR for the COVID-19 outbreak in China using serial interval data --- our analysis shows that using incorrectly defined serial intervals can severely bias the estimate.
This study demonstrates the importance of early epidemiological investigation through contact tracing and provides a rationale for reassessing generation intervals, serial intervals, and \RR\ estimates, for COVID-19.

@@

The reproduction number \RR is one of the most important characteristics of an emerging epidemic, including the current pandemic of coronavirus disease 2019 (COVID-19) \citep{majumder2020early}.
The reproduction number is defined as the average number of secondary cases caused by a primary case;
the value in a fully susceptible population --- referred to as the basic reproduction number \Ro --- allows us to predict the extent to which an infection will spread in the population and the amount of intervention necessary to eliminate it \citep{anderson1991infectious}.
Near the beginning of an outbreak, \RR is often estimated from the observed exponential growth rate using generation- and serial-interval distributions (e.g., \cite{wallinga2007generation, fraser2009pandemic, hampson2009transmission, chunara2012social, chowell2014west, du2020serial, jung2020real}) but the differences in the distributions are often neglected.

@@

The time delays between two epidemiological events can be defined either within an infected individual (e.g., incubation period: infection and symptom onset of an individual) or between infected individuals (e.g., serial interval: symptom onsets of an infector and an infectee).
We can further divide these events into \emph{primary} and \emph{secondary} events.
When we measure an epidemiological time delay within an infected individual (e.g., the incubation period), the primary event is the event that usually occurs before the secondary event ---
most epidemiological events that can be observed within an individual have clear direction (e.g., infection to onset of symptoms) but some may not (e.g., onset of infectiousness and onset of symptoms).
When we measure an epidemiological time delay between infected individuals (e.g., the serial interval), 
the primary and secondary events are defined in terms of the direction of transmission:
the primary event refers to the event that occurs in the infector. 
Again, some of these delays are always positive (the infector is always infected before the infectee) and some are not (it is possible for the infectee to develop symptoms before the infector \citep{he2020temporal}).

@@

Both forward and backward perspectives
must yield identical \emph{measurement} (e.g., the length of the incubation
period of a given individual is the same whether measured forward from
the time of infection or backward from the time of symptom onset).
Consequently, if we restrict attention to individuals with a given
incubation period, say $\tau$ days, and we count how many became infected at time $\psymp$
or how many first showed symptoms at time $\ssymp=\psymp+\tau$ then we must obtain
the same number. In general, no matter how delays are distributed, if
$\mathcal P$ and $\mathcal S$ represent the sizes of primary and
secondary cohorts then we must have
\begin{equation}
\psize(\psymp) f_\psymp(\tau) = \ssize(\ssymp) b_\ssymp(\tau) \,.
\label{eq:match}
\end{equation}
Substituting $\psymp=\ssymp-\tau$, it follows that
\begin{equation}
b_\ssymp(\tau)= \frac{\psize(\ssymp-\tau) f_{\ssymp-\tau} (\tau)}{\ssize(\ssymp)}\,.
\label{eq:backward}
\end{equation}
If we are thinking of realized incubation periods, then the left hand side of
this equation is the probability density that an individual
who became symptomatic at time $\ssymp$ had an incubation period of
length $\tau$. From the right hand side, we see that this probability
density depends on how many individuals became infected at time
$\ssymp-\tau$ [$\psize(\ssymp-\tau)$], how many individuals became
symptomatic at time $\ssymp$ [$\ssize(\ssymp)$], and the probability density
that someone who was infected at time $\ssymp-\tau$ became symptomatic at
time $\ssymp$ [$f_{\ssymp-\tau}(\tau)$]. In general, Eq.~\eqref{eq:backward}
shows that the backward delay distribution at reference time $\ssymp$
[$b_\ssymp(\tau)$] depends not only on the secondary cohort size at time
$\ssymp$ [$\ssize(\ssymp)$], but also on the primary cohort size at the
earlier time $\ssymp-\tau$ [$\psize(\ssymp-\tau)$] and the forward delay
distribution at that earlier time [$f_{\ssymp-\tau}(\tau)$].

Several different mechanisms drive the changes in forward and backward delay distributions over time.
Typically, within-individual forward delay distributions are not affected by epidemic dynamics.
Some realized distributions, like incubation period distributions, are expected to be equivalent to their intrinsic distributions and remain invariant at time time scale of an outbreak.
Other realized distributions, like the distribution of time from symptom onset to testing, may change over the course of an epidemic due to changes in public health policies.
Between-individual forward delay distributions, such as generation- or serial-interval distributions, depend on epidemic dynamics;
for example, forward generation intervals contract during the outbreak because infected individuals are less likely to infect others later in the epidemic due to factors that drive time-dependent decreases in transmission such as susceptible depletion, behavioral change, and interventions \citep{champredon2015intrinsic}.

\eref{backward} suggests that backward delay distributions change over time even if their corresponding forward delay distribution does not change.
Backward delay distributions depend on changes in the primary cohort size over time due to conditionality of observations:
Conditioning on individuals whose secondary events have occurred at the same time means that the distribution is truncated so that we tend to observe shorter (or longer) inter-event delays when incidence is  the cohort size has been increasing (or decreasing) over time.
In particular, when the incidence is growing exponentially, we can calculate the amount of bias exactly.
Assuming that the forward delay distribution remains constant during the exponential growth phase ($f_p(\tau) \approx f_0(\tau)$), we can substitute $\psize(t) = \psize(0) \exp(rt)$ in \eref{backward} to obtain:
\begin{align}
b_0(\tau) &= \exp(-r\tau) f_0(\tau) \left[\psize(0)/\ssize(0)\right]\\
&=\frac{\exp(-r\tau) f_0(\tau)}{\int_{-\infty}^{\infty} \exp(-r\tau') f_0(\tau') \dtau'},
\label{eq:backexp}
\end{align}
where $r$ is the exponential growth rate and the denominator is a constant since it does not depend on $\tau$.
For faster epidemics (higher $r$), the backward delay distribution will have a shorter mean.
Changes in the backward delay distribution are also by definition affected by the changes in the forward delay distribution (\eref{backward}).
Therefore, the backward delay distribution will generally differ from the forward delay distribution even if we are measuring time delays that are intrinsic to the life history of a disease (e.g., the incubation period).
These ideas apply to all epidemiological delay distributions and generalize the work by \cite{champredon2015intrinsic} who compared forward and backward generation-interval distributions to describe the realized generation intervals from the perspective of an infector and an infectee, respectively, as well as the work by \cite{britton2019estimation} who showed that \eref{backexp} holds for the backward generation-interval distributions.

@@

We illustrate changes in the forward and backward serial intervals over the course of an epidemic by applying it to a specific example of an epidemic model. 
We model disease spread with a renewal-equation model \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral, roberts2007model, champredon2018equivalence}.
Ignoring births and deaths, changes in the proportion of susceptible individuals $S(t)$ and incidence of infection $i(t)$ can be written as:
\begin{align}
\frac{\mathrm{d}S}{\mathrm{d}t} &= - i(t)\nonumber\\
i(t) &= \Ro S(t) \int_0^\infty i(t-\tau) \gdist(\tau) \dtau,
\label{eq:renewal}
\end{align}
where \Ro is the basic reproduction number, and $\gdist(\tau)$ is the intrinsic generation-interval distribution (i.e., the forward generation-interval distribution of a primary case in a population where the number of susceptibles remains constant).
Then, the forward generation-interval for a cohort of individuals that were infected at time $\psymp$ follows \citep{champredon2015intrinsic}:
\begin{equation}
\gdist_\psymp (\tau) = \frac{ \gdist(\tau) S(\psymp + \tau)}{\int_0^\infty \gdist(\tau') S(\psymp + \tau') \dtau'},
\end{equation}
which allows us to separate the joint probability distribution $h_\psymp$ of the forward incubation period and the forward generation-interval distribution as a product of the proportion of susceptible individuals $S$ and the joint probability distribution $h$ of the forward incubation period and the intrinsic generation intervals:
\begin{equation}
h_\psymp (x, \tau) = \frac{h(x, \tau) S(\psymp + \tau)}{\int_0^\infty \int_0^\infty h(x', \tau') S(\psymp + \tau') \dtau' \dx'}.
\end{equation}
We further assume that the forward incubation period distribution does
not vary across cohorts over the course of an epidemic, as it
represents the life history of a disease; we denote it as $\idist$.
Then, we have:
\begin{align}
\idist(x) &= \int_0^\infty h(x, \tau) \dtau\nonumber\\
\gdist(\tau) &= \int_0^\infty h(x, \tau) \dx
\label{eq:marginal}
\end{align}
Finally, the case reproduction for this model is defined as follows:
\begin{equation}
\Rc(t) = \Ro \int_0^\infty \gdist(\tau) S(t+\tau) \dtau.
\end{equation}
The forward and backward serial-interval distributions are then calculated by substituting these quantities into \eref{fserial} and \eref{bserial}.
We use this framework to illustrate how the realized epidemiological time distributions vary over the course of an epidemic and depend on the perspective (i.e., forward vs. backward).

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.95\textwidth]{forward.pdf}
\caption{
\textbf{Epidemiological dynamics and changes in mean forward and backward delay distributions.}
(A) Daily incidence over time.
(B--D) Changes in the mean forward incubation period, generation interval, and serial interval.
(E--G) Changes in the mean backward incubation period, generation interval, and serial interval.
Black lines represent the results of a deterministic simulation.
Orange lines (A) represent the results of 10 stochastic simulations.
Orange points (B-G) represent the average of 10 stochastic simulations.
Dashed lines represent the mean initial forward delay.
Intrinsic incubation periods and intrinsic generation-intervals are assumed to be independent of each other.
See Table 1 for parameter values.
}
\label{fig:epi}
\end{center}
\end{figure}

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{serial_analysis.pdf}
\caption{
\textbf{Observed serial intervals of COVID-19 and cohort-averaged estimates of \RR.}
(A--B) forward and backward serial intervals over time.
Points represent the means. 
Vertical error bars represent the 95\% equi-tailed quantiles.
Solid lines represent the estimated locally estimated scatterplot smoothing (LOESS) fits.
The dashed line represents the maximum and minimum observable delays across the range of reported sypmtom onset dates.
(C) Cohort-averaged estimates of \Ro assuming doubling period of 6 and 8 days \citep{li2020early, wu2020nowcasting}.
Ribbons represent the associated 95\% bootstrap confidence intervals.
The data were taken from Supplementary Materials of \cite{du2020serial}.
}
\label{fig:du}
\end{figure}

@@

Since the right hand side of \eref{ptarget} is also a
Laplace transform of $f_0=\hat{z}*\idist$, we can express it as the product
of Laplace transforms of $\hat{z}$ and $\idist$:
\begin{equation}
\int_{-\infty}^\infty \exp(-r\tau) f_{0}(\tau) \mathrm{d} \tau = \int_{-\infty}^\infty \exp(-r\tau) \hat{z}(\tau)\, \mathrm{d} \tau \int_{0}^\infty \exp(-r\tau) \idist(\tau) \dtau\,.
\label{eq:newtarget}
\end{equation}
In order to derive an expression for a Laplace transform of $\hat{z}$, we have to first derive an analytical expression for $\int_{-\infty}^\infty z(x) \dx$. By changing the order of integration, we have:
\begin{align}
\int_{-\infty}^\infty z(\sinf) \mathrm{d}\sinf &= \int_{-\infty}^\infty \int_{\max{(0,-\sinf)}}^{\infty} \exp(-r\pinf) h(\pinf, \sinf + \pinf)\, \mathrm{d}\pinf \,\mathrm{d}\sinf\nonumber\,,\\
&= \int_{0}^\infty \int_{-\pinf}^\infty \exp(- r \pinf) h(\pinf, \sinf+\pinf)\, \mathrm{d}\sinf\,\mathrm{d} \pinf\,.
\end{align}
Since $\idist$ is a marginal probability distribution of $h$, it follows that:
\begin{equation}
\int_{-\infty}^\infty z(\sinf) \mathrm{d}\sinf = \int_{0}^\infty \exp(- r \pinf) \idist(\pinf)\, \mathrm{d}\pinf\,.
\end{equation}
Then, we have:
\begin{equation}
\hat{z}(\sinf) = \frac{\int_{\max{(0,-\sinf)}}^{\infty} \exp(-r\pinf) h(\pinf, \sinf + \pinf)\, \mathrm{d}\pinf}{\int_{0}^\infty \exp(- r \pinf) \idist(\pinf)\, \mathrm{d}\pinf}\,.
\end{equation}
Substituting the expression into \eref{newtarget}, we have:
\begin{equation}
\int_{-\infty}^\infty \exp(-r\sinf) f_{0}(\tau)\dtau = \int_{-\infty}^\infty \exp(-r\sinf) \int_{\max{(0,-\tau)}}^{\infty} \exp(-r\pinf) h(\pinf, \tau + \pinf)\, \mathrm{d}\pinf\,\mathrm{d}\sinf\,.
\label{eq:newtarget2}
\end{equation}

Recall that $g$ is also a marginal probability distribution of $h$:
\begin{equation}
g(\tau) = \int_0^\infty h(x, \tau) \,\mathrm{d}x\,.
\end{equation}
We can then substitute $\tau = \pinf + \sinf$ into \eref{newtarget2} and apply change of variables to obtain:
\begin{align}
&\int_{-\infty}^\infty \exp(-r\tau) f_{0}(\tau)\dtau\\
&=\int_{-\infty}^{\infty} \exp(-r\sinf) \int_{\max(0, -\sinf)}^\infty \exp(- r \pinf) h(\pinf, \tau+\pinf) \,\mathrm{d} \pinf \,\mathrm{d}\sinf \\
&=\int_{0}^{\infty} \int_{0}^\infty \exp(- r \tau) h(\pinf, \tau)\, \mathrm{d} \pinf\, \mathrm{d}\tau\\
&=\int_{0}^{\infty} \exp(-r\tau) g(\tau) \dtau =\frac{1}{\Ro}
\end{align}
Therefore, the initial forward serial-interval distribution and the intrinsic generation-interval distribution give the same estimates of \Ro from $r$.\qed

### 2020-05-28 serial_analysis.tex
tweaks
almost done

% Created by tikzDevice version 0.12.3 on 2020-05-14 21:20:58
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

@@

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (206.43,192.30) {maximum observable};

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 92.13, 70.86) {minimum observable};

@@

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (388.02,200.87) {maximum observable};

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (521.60, 75.14) {minimum observable};

>>>>

% Created by tikzDevice version 0.12.3 on 2020-05-27 22:31:58
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

@@

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (210.00,195.87) {maximum observable};

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 95.70, 74.43) {minimum observable};

@@

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (384.09,204.80) {maximum observable};

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (517.67, 79.07) {minimum observable};
